---
title: "Previous_Time_Series_Work"
author: "Mitchell Nelson"
date: "2023-05-06"
output: html_document
---


## Libraries

```{r}
library(readr)
library(ggplot2)
```


### Simple linear trend mapped onto sample data with predictions for the next 5 data points and a 95% confidence interval graphed.

```{r}
fin1_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final1-data.txt", col_names = FALSE)
fin1_x <- fin1_dat$X1
findex1 <- c(1:50)
fin1_dat <- as.data.frame(cbind(fin1_x, findex1))

n <- length(fin1_x)
a_hat_post <- sum(fin1_x)/n
t_bar <- (n+1)/2

t <- findex1 - t_bar
bot_comp <- sum(t^2)
top_comp <- sum((fin1_x - a_hat_post) * t)
b_hat <- (top_comp)/(bot_comp)
int <- a_hat_post - (b_hat * 25.5)
int
vare <- (sum((fin1_x - a_hat_post - b_hat*t)^2))/(n-2)
sqvare <- sqrt(vare)
t_stat <- (b_hat*sqrt(bot_comp))/sqvare

#t_stat
t_crit <- abs(qt(p = 0.025, df = 49))

trend <- function(m){
  res <- a_hat_post + b_hat*(n+m-t_bar)
  return(res)
}
lower <- function(m){
  partres <- t_crit * sqvare * sqrt(1 + 1/n + ((n+m-t_bar)^2)/bot_comp)
  res <- trend(m) - partres
  return(res)
}
  
upper <- function(m){
  partres <- t_crit * sqvare * sqrt(1 + 1/n + ((n+m-t_bar)^2)/bot_comp)
  res <- trend(m) + partres
  return(res)
}

trend(1)
lower(1)
upper(1)

low <- c(rep(0,5))
fit <- c(rep(0,5))
upp <- c(rep(0,5))

for(i in 1:5){
  low[i] <- lower(i)
  fit[i] <- trend(i)
  upp[i] <- upper(i)
}

ext_index1 <- c(51:55)

full_dat1 <- cbind(low, fit)
full_dat1 <- cbind(full_dat1, upp)
full_dat1 <- as.data.frame(cbind(full_dat1, ext_index1))
full_dat1

ggplot(data = fin1_dat, aes(x = findex1, y = fin1_x)) + geom_point() + coord_cartesian(xlim = c(0, 56)) + geom_abline(aes(intercept = int, slope = b_hat)) + geom_line(data = full_dat1, aes(x = c(51:55), y = low), color = "red") + geom_line(data = full_dat1, aes(x = c(51:55), y = upp), color = "red") + labs(x = "Time", y= "Outcome")
```


### Application of "53X" smoothing method to sample data


```{r}
fin2_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final2-data.txt", col_names =  FALSE)
fin2_x <- fin2_dat$X1
length(fin2_x)
findex2 <- c(1:85)
fin2_dat <- as.data.frame(cbind(findex2, fin2_x))


med1 <- c()
med2 <- c()
fin <- c()


for(s in 3:85){
  s1vec <- c(fin2_x[s-2], fin2_x[s-1], fin2_x[s], fin2_x[s+1], fin2_x[s+2])
  s1 <- median(s1vec)
  med1 <- c(med1, s1)
}

for(h in 2:82){
  s2vec <- c(med1[h-1], med1[h], med1[h+1])
  s2 <- median(s2vec)
  med2 <- c(med2, s2)
}

for(b in 2: 80){
  s3 <- (0.25 * med2[b-1]) + (0.5 * med2[b]) + (0.25 * med2[b+2])
  fin <- c(fin, s3)
}

plot(fin2_x)
plot(fin)
```


### Exponential Smoothing applied to sample data - 0th order

```{r}
fin3_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final3-data.txt", col_names = FALSE)
fin3_x <- fin3_dat$X1
length(fin3_x)
findex3 <- c(1:80)
fin3_dat <- as.data.frame(cbind(findex3, fin3_x))
plot(fin3_x)

alpha <- 0.1
ahati <- c(11.765550)

## 0-order

for(h in 2:80){
  pe <- fin3_x[h] - ahati[h-1]
  newahat  <- (pe * alpha) + ahati[h-1]
  ahati <- c(ahati, newahat)
}

preds <- c(rep(ahati[80], 5))


preds

z_exp <- cbind(findex3, fin3_x)
z_exp <- as.data.frame(cbind(z_exp, ahati))
z_exp

ggplot(data = z_exp, aes(x = findex3, y = ahati)) + geom_point() + coord_cartesian(xlim = c(0, 87)) + geom_point(data = pred_dat, aes(x = c(81:85), y = preds), color = "red")
```

### Exponential Smoothing applied to sample data - 1st order

```{r}

ahati1 <- c(11.765550)
bhati1 <- c(0)
pred1vec <- c()
pred1 <- 0
 
for(j in 2:81){
  pred1 <- ahati1[j-1] + bhati1[j-1]
  pred1vec <- c(pred1vec, pred1)
  pe1 <- fin3_x[j] - pred1
  newahat1 <- ahati1[j-1] + bhati1[j-1] + (2*alpha + (alpha)^2) * pe1 
  newbhat1 <- bhati1[j-1] + (alpha)^2 * pe1
  ahati1 <- c(ahati1, newahat1)
  bhati1 <- c(bhati1, newbhat1)
}
o_exp <- cbind(findex3, ahati1)
o_exp <- cbind(o_exp, bhati1)
o_exp <- as.data.frame(cbind(o_exp, pred1vec))
o_exp <- o_exp[-81,]
o_exp

preds1 <- c()
for(i in 1:5){
  calc <- ahati1[80] + i * bhati1[80]
  preds1 <- c(preds1, calc)
}

ext_index3 <- c(81:85)
pred_dat1 <- as.data.frame(cbind(ext_index3, preds1))
pred_dat1

ggplot(data =  o_exp, aes(x = findex3, y = pred1vec)) + geom_point() + coord_cartesian(xlim = c(0, 87)) + geom_point(data = pred_dat1, aes(x = c(81:85), y = preds1), color = "red")
```
### Exponential Smoothing applied to sample data - 2nd order

```{r}
ahati2 <- c(11.765550)
bhati2 <- c(0)
chati2 <- c(0)
pred2vec <- c()

for(y in 2:81){
  pred2 <- ahati2[y-1] + bhati2[y-1] + chati2[y-1]
  pred2vec <- c(pred2vec, pred2) 
  pe2 <- fin3_x[y] - pred2
  newahat2 <- ahati2[y-1] + bhati2[y-1] + chati2[y-1] + (alpha * (3 - (3 * alpha) + alpha^2)) * pe2
  newbhat2 <- bhati2[y-1] + 2 * chati2[y-1] + (3 * alpha^2 * (1 - (alpha/2))) * pe2
  newchat2 <- chati2[y-1] + ((alpha^3)/2) * pe2
  ahati2 <- c(ahati2, newahat2)
  bhati2 <- c(bhati2, newbhat2)
  chati2 <- c(chati2, newchat2)
}


t_exp <- cbind(findex3, ahati2)
t_exp <- cbind(t_exp, bhati2)
t_exp <- cbind(t_exp, chati2)
t_exp <- as.data.frame(cbind(t_exp, pred2vec))
t_exp <- t_exp[-81,]
t_exp

preds2 <- c()
for(i in 1:5){
  calc <- (ahati2[80] + (i * bhati2[80])  + (chati2[80] * i^2))
  calc
  preds2 <- c(preds2, calc)
}
ext_index3 <- c(81:85)
pred_dat2 <- as.data.frame(cbind(ext_index3,preds2))
pred_dat2


ggplot(data = t_exp, aes(x = findex3, y = pred2vec)) + geom_point() + coord_cartesian(xlim = c(0, 87)) + geom_point(data = pred_dat2, aes(x = ext_index3, y = preds2), color = "red")
```




### Data formatting

```{r}
fin4_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final4-data.txt", col_names = FALSE)
fin4_x <- fin4_dat$X1

```

### Hand-coded Kendall Rank Correlation Test applied to sample data

```{r}
K <- 2  #pairs variable
n <- length(fin4_x)

for ( i in 1:n){
  for (j in 2:n){
   if (fin4_x[i] < fin4_x[j] && i < j){
     K <- K + 1
   }
  }
}

K

T <- ((4 * K)/(n * (n-1))) -1
SDT <- sqrt((2 * ((2 * n) + 5))/((9 * n) * (n -1)))

Tstat <- T/SDT
Tstat
```

### Hand-coded Spearman Rank Correlation Test applied to sample data

```{r}
sort4 <- c(sort(fin4_x))
ordered4 <- order(fin4_x)

findex4 <- 1:50

ranks <- c(rank(fin4_x))
fin4_sorting <- cbind(sort4, fin4_x, ranks, findex4)
fin4_sorting


```


```{r}
rankres <- c(1:50)
for (i in 1:50){
  rankres[i] <- (ranks[i] - i)^2
}
```

```{r}
S <- 1 - (6 * sum(rankres))/(n * (n^2-1))
S
SSD <- sqrt(1/n)
SSD
SStat <- S/SSD
SStat
```
#### As both tests exceed 1.96 in absolute value, we reject the null of no trend for both.




### Estimation of mean, variance, 1st 25 Values of ACF, 1st 3 Values of PACF and Spectral Density with a triangular window and M=25.

```{r}
fin8_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final8-data.txt", col_names = FALSE)
fin8_x <- fin8_dat$X1
fin8_dat
length(fin8_x)
```

## Mean and Var
```{r}
x1bar <- mean(fin8_x)
x1bar
  
n <- length(fin8_x)

p1 <- (fin8_x - x1bar)^2
p2 <- (fin8_x - x1bar)


sigma1hatsquare <- (sum(p1))/length(fin8_x)
sigma1hatsquare
```
## ACV Calc
```{r}
  k_sum <- c()
acv_est <- function(k){
  k_sum <- c()
for(i in 1:(n-k)){
  res <- (fin8_x[i] - x1bar) * (fin8_x[i+k] - x1bar)
  k_sum <- c(k_sum, res)
}
  full <- sum(k_sum)/n
  return(full)
}

acv_vals <- c(1:26)
acv_vals[1] <- sigma1hatsquare
for(i in 1:25){
  acv_vals[i+1] <- acv_est(i)
}

acv_dat <- as.data.frame(cbind(c(0:25), acv_vals))
acv_dat

plot(acv_vals)
acf(fin8_x, type = "covariance")
```
## ACF Calc
```{r}
acf_vals <- c(rep(0,26))
acf_vals[1] <- 1
for(i in 1:26){
  acf_vals[i] <- acv_vals[i]/ sigma1hatsquare
}
plot(acf_vals)
acf_dat <- as.data.frame(cbind(c(0:25), acf_vals))
acf_dat
acf(fin8_x, type = "correlation")
```

```{r}
pacf_vals <- c(1:4)
pacf_vals[1] <- 1
pacf_vals[2] <- acf_vals[2]
v1 <- c(1, acf_vals[2],acf_vals[2], acf_vals[3])
v2 <- c(1, acf_vals[2],acf_vals[2],1)
v12 <- c(1, acf_vals[2],acf_vals[2], acf_vals[2], 1, acf_vals[3], acf_vals[3], acf_vals[2], acf_vals[4])
v22 <- c(1, acf_vals[2],acf_vals[3], acf_vals[2], 1, acf_vals[2], acf_vals[3], acf_vals[2], 1)
mat1 <- matrix(v1, nrow = 2)
mat2 <- matrix(v2, nrow = 2)
mat12 <- matrix(v12, nrow = 3)
mat22 <- matrix(v22, nrow = 3)

pacf_vals[3] <- det(mat1)/det(mat2)
pacf_vals[4] <- det(mat12)/det(mat22)

pacf_dat <- cbind(c(0:3), pacf_vals)
pacf_dat
acf(fin8_x, type = "partial")
```

```{r}
M <- 25
om <- seq(from = 0, to = pi, by = 0.001)
res <- ((1-(1/M)) * cos(1*om)*acv_vals[2]) + ((1-(2/M)) * cos(2*om)*acv_vals[3]) + ((1-(3/M)) * cos(3*om)*acv_vals[4]) + ((1-(4/M)) * cos(4*om)*acv_vals[5]) + ((1-(5/M)) * cos(5*om)*acv_vals[6]) + ((1-(6/M)) * cos(6*om)*acv_vals[7]) + ((1-(7/M)) * cos(7*om)*acv_vals[8]) + ((1-(8/M)) * cos(8*om)*acv_vals[9]) + ((1-(9/M)) * cos(9*om)*acv_vals[10]) + ((1-(10/M)) * cos(10*om)*acv_vals[11]) + ((1-(11/M)) * cos(11*om)*acv_vals[12]) + ((1-(12/M)) * cos(12*om)*acv_vals[13] )+ ((1-(13/M)) * cos(13*om)*acv_vals[14]) + ((1-(15/M)) * cos(15*om)*acv_vals[16]) + ((1-(16/M)) * cos(16*om)*acv_vals[17]) + ((1-(17/M)) * cos(17*om)*acv_vals[18]) + ((1-(18/M)) * cos(18*om)*acv_vals[19]) + ((1-(19/M)) * cos(19*om)*acv_vals[20]) + ((1-(20/M)) * cos(20*om)*acv_vals[21]) + ((1-(21/M)) * cos(21*om)*acv_vals[22]) + ((1-(22/M)) * cos(22*om)*acv_vals[23]) + ((1-(23/M)) * cos(23*om)*acv_vals[24]) + ((1-(24/M)) * cos(24*om)*acv_vals[25]) + ((1-(25/M)) * cos(25*om)*acv_vals[26])  

tri_wind <- (1/(2*pi)) * (acv_vals[1] + (2 * res))
plot(tri_wind, type = "l")
```



### Transformation and graph of transfer function for the below filter.

#### The provided equation of (1.1501Yt - 0.621438Yt-1 + 0.894896Yt-2 = Xt - 0.621438Xt-1 + Xt-2) is equivalent to (Yt -0.56233644Yt-1  + 0.80978734Yt-2 = 0.90489548(Xt -0.621438Xt-1 + Xt-2)).  With this transformation, we can graph it as follows.

```{r}
w <- seq(from = 0, to = pi, by = 0.01)

a1 <- -0.56233644
a2 <- 0.80978734
b1 <- -0.621438
b2 <- 1

alpha <- 1 + a1^2 + a2^2 + 2*a1*(1+a2)*cos(w) + 2*a2*cos(2*w)
beta <- 1 + b1^2 + b2^2 + 2*b1*(1+b2)*cos(w) + 2*b2*cos(2*w)

graph <- (0.90489548*(beta)/alpha)
plot(graph, type ="l")
```
#### This filter is a band-reject filter with cutoffs around 1.1 and 1.3.

### Application of a 2nd Order Band Pass Tangent Filter with cutoff of 8pi/11 +- pi/30 to sample data

```{r}
fin10_dat <- read_csv("C:/Users/mitch/Documents/MINE/R_CODING/RCODE/DATASETS/23-Final10-data.txt", col_names = FALSE)
fin10_x <- fin10_dat$X1
fin10_y <- c(rep(0,1350))
```

```{r}
D <- cos((8*pi)/11)/cos(pi/30)
E <- tan(pi/30)

for(i in 3:1350){
  partres =  E * (fin10_x[i] - fin10_x[i-2]) + (2 * D * fin10_y[i-1]) - ((1-E) * fin10_y[i-2])
  fin10_y[i] = partres / (E+1)
}

findex10 <- c(1:1350)
head(fin10_y)

full_10_dat <- cbind(findex10, fin10_y)

ggplot(data = as.data.frame(full_10_dat), aes(x = findex10, y = fin10_y)) + geom_point() 

```