---
title: "AQ13_ISLR"
author: "Mitchell Nelson"
date: "2022-09-18"
output: html_document
---

# Applied Question 13 from ISLR Book
## Modeling and predicting stock data based on various weeks lag using various statistical classification methods including: 
## Regular Logistic Regression, Linear Discriminant Analysis, Quadratic Discriminant Analysis, K-Nearest Neighbors, and Naive Bayes methods

```{r}
library(ISLR2)
library(MASS)
library(class)
library(e1071)
```

## Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?
```{r}
head(Weekly)
plot(Weekly)
Weekly
```


#### Within our data, there are a couple of patterns present in the correlation plot (derived from the plot function).  Notably, we see distinct correlation between Year and VOlume which is a positive exponential model. As far as the other variables are concerned, there does not seem to be any other stand out patterns.

## Creating Logistic Regression model to see significant variables

```{r}
glm.full <- glm(
    Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
    data = Weekly, family = binomial
  )
summary(glm.full)
```

####  After running our logistic regression for all of our features against direction, we can see that Lag2 is the only significant feature within our model.


## Confusion matrix for our logistic regression

```{r}
glm.probs <- predict(glm.full, type = "response")  
glm.probs[1:10]
glm.pred <- rep("Down", 1089)
glm.pred[glm.probs > .5] = "Up"
table(glm.pred, Weekly$Direction)
mean(glm.pred == Weekly$Direction)
```

#### From our confusion matrix above, we can see that our logistic regression over-predicts the amount of times that the market will go up with 430 mispredictions.  On average our model is correct roughly 56% of the time.

## Fitting Logistic Regression model using training data from 1990-2008 and testing after

```{r}
Weekly.tr <- Weekly[Weekly$Year >= 1990 & Weekly$Year <= 2008,]
Weekly.ts <- Weekly[Weekly$Year > 2008,]
glm.red <- glm(
    Direction ~ Lag2,
    data = Weekly.tr, family = binomial
  )
summary(glm.red)
dim(Weekly.tr)
```

## Computing confusion matrix against testing data

```{r}
glm.red.probs <- predict(glm.red, Weekly.ts, type = "response")  
glm.red.pred <- rep("Down", 104)
glm.red.pred[glm.red.probs > .5] = "Up"
table(glm.red.pred, Weekly.ts$Direction)
mean(glm.red.pred == Weekly.ts$Direction)
```

## Fitting a Linear Discriminant Analysis similarly to our logistic regression 

```{r}
lda.fit <- lda(Direction ~ Lag2,
    data = Weekly.tr)
lda.fit
plot(lda.fit)
```

## Confusion matrix for testing data

```{r}
lda.pred <- predict(lda.fit, Weekly.ts)
lda.class <- lda.pred$class
table(lda.class, Weekly.ts$Direction)
mean(lda.class == Weekly.ts$Direction)
```

## Fitting a Quadratic Discriminant Analysis

```{r}
qda.fit <- qda(Direction ~ Lag2,
    data = Weekly.tr)
qda.fit
```
## Confusion matrix for testing data

```{r}
qda.class <- predict(qda.fit, Weekly.ts)$class
table(qda.class, Weekly.ts$Direction)
mean(qda.class == Weekly.ts$Direction)
```
### Weird result of only predicting that the market will go up, if only!

## Fitting KNN model to our training data 

## First need to create matrices for our function

```{r}
train <- (Weekly$Year >= 1990 & Weekly$Year <= 2008)
train.X <- cbind(Weekly$Lag2[train])
test.X <- cbind(Weekly$Lag2[!train])
train.Dir <- Weekly$Direction[train]

```


```{r}
set.seed(1)
knn.pred <- knn(train.X, test.X, train.Dir, k = 1)
table(knn.pred, Weekly.ts$Direction)
mean(knn.pred == Weekly.ts$Direction)
```
## Fitting a Naive Bayes model


```{r}
nb.fit <- naiveBayes(Direction ~ Lag2, data = Weekly.tr)
nb.fit

```

## Confusion Matrix for testing data

```{r}
nb.class <- predict(nb.fit, Weekly.ts)
table(nb.class, Weekly.ts$Direction)
mean(nb.class == Weekly.ts$Direction)

```








